---
title: "It's the Economy Stupid: A Time Series Analysis of 10 Year Daily Treasury Rates with ANOVA"
author: "Rebecca Rodriguez, Erik Daenitz, Nick Hickman, Seoyeon Yang"
date: "April 22, 2018"
output: html_document
---


## Step 1: Data prep
Stream U.S. 10 yeare Treasury data into data frame. We use daily end-of-day observations for yields. We then subset that data to take the most recent ~20 years of daily observations.

```{r, tidy = TRUE}
## Stream data into dataframe
treasuryRates = read.csv("./FED-SVENPY.csv") #daily rates

## Subset dataframe to only include 10 year rates
treasury10Y = subset(treasuryRates, select=c(Date,SVENPY10))
colnames(treasury10Y)[c(1,2)] = c("Date", "Value") # rename columns
treasury10Y = treasury10Y[complete.cases(treasury10Y), ] # drop NAs
treasury10Y = treasury10Y[(order(as.Date(treasury10Y$Date))),]
row.names(treasury10Y) = NULL  # re-index the new subset
rates = treasury10Y[((nrow(treasury10Y))-((252*30)+1)):(nrow(treasury10Y)),] # take ~30 past years of treasury data, data takes a year to = 252 days per year and 30 years of past data from the latest observed value
rownames(rates) = NULL # re-index the new subset

# Plot time series for fun visualization
plot(rates$Value ~ as.Date(rates$Date), type="l", main = "U.S. Treasury 10 Year - Yield (%)", xlab = "Date" ,ylab = "Yield (%)")

```

## Step 2: Use Monte Carlo Method to simulate rates

1) Generate X random standard normal variables (X = number of observations over 20 yr period)
2) Repeat 1) 1000 times to obtain 7562x1000 matrix
3) Perform cascading One Way ANOVA test within each matrix column
  a) Perform X-4 tests and obtain X-4 test statistics
  b) Grab the max test-statistic for each column to obtain 1000 max values to form our Lambda Distribution with

```{r, tidy = TRUE}
set.seed(1000) # Set seed to ensure replicable results

# Parameters for sampling from standard normal distribution
n = length(rates$Value) # The sample size
sims = 1000 # The number of simulations to replicate, also the length of discrete values in our Lambda Distribution
mu = 0 # The true expected mean value of our normal distribution
s = 1 # The true variance of our normal distribution

# Generate random samples of size n
rs = matrix(rnorm(n*sims,mu,s), nrow=n, ncol=sims)

# Function to perform cascading ANOVA tests through each vector and return max F value
# rs: vector of float values
anova_tests = function(rs){
  index_end = length(rs)
  #vectors to hold test statistics
  anovatest_stats = c()
  #traverse through list and perform anova tests
  for (i in 2:(index_end-2)){
    #create factors for anova test
    populations = factor(c(rep(c("a","b"), c(i,index_end-i))))
    #perform anova test
    anova_results = unlist(summary(aov(rs~populations)))
    #put test statistic in list
    anovatest_stats = c(anovatest_stats, anova_results[7])
  }
  #return max test statistic
  return(max(anovatest_stats))
}

## Now run the function over every vector in the matrix ##

# Load the library "snow"" to parallelize the computation using all available cores of the local machine
library(snow)
# Open up sockets to the 4 local processor cores
## Note: only enter "localhost" = to the number of cores on your processor
cl = makeSOCKcluster(c("localhost","localhost","localhost","localhost"))

# Retrieve max test statistics by applying the function to each matrix vector
## Note: expect to wait about 45 mins on a 2.7ghz quad core processor ##
max_stats_anova2 = parApply(cl, rs, 2, anova_tests)
```


##Step 3: Create Lambda distribution CDF, PDF and critical values from simulated Lamda Distribution
Now that we have achieved the simulation of the Lamda Distribution we must assemble a CDF and take quantiles of the PDF at different alpha critical value levels.

Our chosen alpha levels:
$$\alpha=0.10, 0.05, 0.025, 0.01, 0.001$$

```{r, tidy = TRUE}
# CDF and plotted PDF using max values Lamda Distribution from ANOVA tests
lambdaCDF_anova = ecdf(max_stats_anova2)
plot(density(max_stats_anova2), main = "Lambda PDF", xlab = "Critical Value" , ylab = "Density of Observations")
plot(lambdaCDF_anova, main = "Lambda CDF", xlab = "Critical Value" , ylab = "Accumulated Area")

# Spare P-value function to create P-values from our underlying Lambda Distribution at any specified critical value
## Note: This function is not used in subsequent code, it is just created to provide quick intuition on associated p-values with any critical value if the user wishes to get a better sense of the distribution areas.
lambdaPVal = function (x){
  return(1-lambdaCDF(x))
} 

# Return appropriate critical values of our Lambda Distribution to later test our F-Statistics on live data against 
lambdaQTSd = quantile(max_stats_anova2, probs = c(0.90, 0.95, 0.975, 0.99, 0.999), na.rm = FALSE, names = FALSE, type = 1)
```

##Step 4: Perform ANOVA tests on treasury data

```{r, tidy = TRUE}
# Function to run on live data that collects max f-values and corresponding index
anova_test_live = function(rs){
  index_end = length(rs)
  #vectors to hold test statistics
  anovatest_stats = c()
  #traverse through list and perform anova tests
  for (i in 2:(index_end-2)){
    #create factors for anova test
    populations = factor(c(rep(c("a","b"), c(i,index_end-i))))
    #perform anova test
    anova_results = unlist(summary(aov(rs~populations)))
    #put test statistic in list
    anovatest_stats = c(anovatest_stats, anova_results[7])
  }
  
  #return vector of max test statistic and corresponding index
  max_stat = max(anovatest_stats)
  return(c(max_stat, (match(max_stat, anovatest_stats)+1))) #index of last element in first population
}

#find biggest changepoint in data
max1 = anova_test_live(rates$Value)

#create two subsets from max1
pop1a = rates$Value[(1:max1[2])]
pop1b = rates$Value[(max1[2]+1):(length(rates_m$Value))]

#perform anova tests on each subset and grab max test statistics
max2 = anova_test_live(pop1a)
max3 = anova_test_live(pop1b)

#create two subsets from max2
pop2a = pop1a[1:max2[2]]
pop2b = pop1a[(max2[2]+1):length(pop1a)]

#create two subsets from max3
pop3a = pop1b[1:max3[2]]
pop3b = pop1b[(max3[2]+1):length(pop1b)]

#perform anova tests on each new subset and grab max test statisticss
max4 = anova_test_live(pop2a)
max5 = anova_test_live(pop2b)
max6 = anova_test_live(pop3a)
max7 = anova_test_live(pop3b)

# Retrieve index from original data
# for example, grab index within original data for max5
# which(rates_m$Value == pop3b[max7[2]])

### Our First Change Point ####
firstChangePoint = rates[max1[2],]
firstChangePValue = lambdaPVal(max1[1])

changePointIndices = c(max1[2],(which(rates$Value == pop1a[max2[2]])),(which(rates$Value == pop1b[max3[2]])),(which(rates$Value == pop2a[max4[2]])),(which(rates$Value == pop2b[max5[2]])),(which(rates$Value == pop3a[max6[2]])),(which(rates$Value == pop3b[max7[2]])))


# Function to perform anova tests returning test statistic on the real data
# rs:vector of float values
anova_F = function(rs){
  index_end = length(rs)
  #vectors to hold test statistics
  anovatest_stats = c()
  #traverse through list and perform anova tests
  for (i in 2:(index_end-2)){
    #create factors for anova test
    populations = factor(c(rep(c("a","b"), c(i,index_end-i))))
    #perform anova test
    anova_results = unlist(summary(aov(rs~populations)))
    f_stat = anova_results[7]
    #get critical val
    c_val = lambdaQTSd[4] # Reference Lambda critical value at alpha = 0.01
    #compare test stat to critical val
    if (f_stat > c_val){ 
      anovatest_stats = c(anovatest_stats, "Changepoint", anova_results[7], anova_results[9])
    }
    #otherwise, no change found
    else{
      anovatest_stats = c(anovatest_stats, "No Changepoint", anova_results[7], anova_results[9])
      }
  }
  #return max test statistic
  return(anovatest_stats)
}

#### Run the test on live data comparing to our Lambda Distribution critical value ###
# We also contruct the steps below to kick out the F-value and P-value for each test run
a_Results = as.data.frame(anova_F(rates$Value))
anova_results = data.frame(
  x = a_Results,
  result = c("Result", "FVal", "PVal")[c(1,2,3)]
)
# Unstack into a dataframe to see each respective index, F-Value and P-Value
anova_results = unstack(anova_results)

max(anova_results$FVal)

par(mfrow=c(2,2))
plot(rates$Value ~ as.Date(rates$Date), type="l", main = "U.S. Treasury 10 Year - Yield (%)", xlab = "Date" , ylab = "Yield (%)")
plot(anova_results$FVal, type="l", main = "F-Values Across Cascading Test", xlab = "Index", ylab = "F-Values")
plot(anova_results$PVal, type = "l", main = "P-Values Across Cascading Test", xlab = "Index", ylab = "P-Values")

###########################

## Subsetting code to come

sub2 = rates[3:length(rates$Value),]

a2_Results_sub2 = as.data.frame(anova_F(sub2$Value))
anova_results_sub2 = data.frame(
  x = a2_Results_sub2,
  result = c("Result", "FVal", "PVal")[c(1,2,3)]
)

anova_results_sub2 = unstack(anova_results_sub2)

plot(sub2$Value ~ as.Date(sub2$Date), type="l", main = "U.S. Treasury 10 Year - Yield (%)", xlab = "Date" , ylab = "Yield (%)")
plot(anova_results_sub2$FVal, type="l", main = "F-Values Across Cascading Test", xlab = "Index", ylab = "F-Values")
plot(anova_results_sub2$PVal, type = "l", main = "P-Values Across Cascading Test", xlab = "Index", ylab = "P-Values")

## Data for plot
aov_chart = rates[-3,]
aov_chart = rates[1:(n-3),]
aov_chart$Date = as.Date(aov_chart$Date)
aov_chart$Value = as.numeric(aov_chart$Value)
aov_chart$FVal = as.numeric(anova_results$FVal)
aov_chart$PVal = as.numeric(anova_results$PVal)
aov_chart$FVal = round(aov_chart$FVal, digits = 3)
aov_chart$PVal = round(aov_chart$PVal, digits = 6)


library(ggplot2)
library(reshape2)
aov_chart_prep = melt(aov_chart, id = "Date", measure = c("Value", "FVal", "PVal"))
ggplot(aov_chart_prep, aes(Date, value, colour = variable)) + geom_line()


chart = ggplot() + 
  geom_line(data = aov_chart, aes(x = Date, y = Value), color = "red") + scale_y_continuous
  geom_line(data = aov_chart, aes(x = Date, y = FVal), color = "blue") + scale_y_continuous(sec.axis = sec_axis(~./1000)) +
  geom_line(data = aov_chart, aes(x = Date, y = PVal), color = "green") +
  xlab('Date')
chart

### Running new data in to come
treasury10Y_new = read.csv("./10YNew.csv") #daily rates
treasury10Y_new = rbind(rates,treasury10Y_new)



```