---
title: "Math 3080 Final Project"
author: "Rebecca Rodriguez, Erik Daenitz, Nick Hickman, Seoyeon Yang"
date: "April 27, 2018"
output: html_document
---


## Step 1: Data prep
Stream U.S. 10 year Treasury data into data frame. We use monthly observations for yields. We then subset that data to take the most recent ~20 years of daily observations.

```{r, tidy = TRUE}

#import monthly data
treasuryRates = read.csv("10YTreasuryYield.csv") #monthly rates

#grab past 20 years
rates = treasuryRates[1:245,]

#remove factor from date column
rates$Date = as.vector(rates$Date)

#convert strings to dates
rates$Date = as.Date(rates$Date, "%m/%d/%Y")

#visualize data
plot(rates$Date, rates$Value, type="l")
```


## Step 2: Use Monte Carlo Method to simulate rates

1) Generate X random standard normal variables (X = number of observations over 20 yr period)
2) Repeat 1) 1000 times to obtain matrix of values
3) Perform cascading One Way ANOVA test within each matrix column
  a) Perform X-4 tests and obtain X-4 test statistics
  b) Grab the max test statistic for each column to obtain 1000 max values to form our Lambda Distribution with


```{r, tidy = TRUE}
set.seed(1000) # Set seed to ensure replicable results

# Parameters for sampling from standard normal distribution
n = length(rates$Value) # The sample size
m = 1000 # The number of simulations to replicate, also the length of values in our Lambda Distribution
mu = 0 # The true expected mean value of our normal distribution
s = 1 # The true variance of our normal distribution

# Generate m random samples of size n
rs = matrix(rnorm(n*m,mu,s), nrow=n, ncol=m)

# Function to perform cascading ANOVA tests through a vector and return max f statistic
# rs: vector of float values
anova_tests <- function(rs){
  index_end = length(rs)
  #vectors to hold test statistics
  anovatest_stats = c()
  #traverse through list and perform anova tests
  for (i in 2:(index_end-2)){
    #create factors for anova test
    populations = factor(c(rep(c("a","b"), c(i,index_end-i))))
    #perform anova test
    anova_results = unlist(summary(aov(rs~populations)))
    #put test statistic in list
    anovatest_stats = c(anovatest_stats, anova_results[7])
  }
  #return vector of max test statistic and corresponding index
  max_stat = max(anovatest_stats)
  return(cbind(max_stat, (match(max_stat, anovatest_stats)+1))) #index of last element in first population
}

## Function to perform t tests
# values: vector of float values
t_tests <- function(values){
  index_end = length(values)
  #vectors to hold test statistics
  ttest_stats = c()
  #traverse through list and perform t tests
  for (i in 2:(index_end-2)){
    #perform t test
    ttest_results = t.test(c(values[1:i]), c(values[i+1:index_end]), alternative="two.sided", var.equal=TRUE)
    #add test statistic to list
    ttest_stats = c(ttest_stats, ttest_results$statistic)
  }
  
  #return vector of (absolute value of) max test statistic and corresponding index
  max_stat = max(abs(ttest_stats))
  if(max_stat > max(ttest_stats)){ #if abs val of max is bigger than max
    max_stat_index = (match(-max_stat, ttest_stats)+1)  #then return index of neg max
  }
  else{
    max_stat_index = (match(max_stat, ttest_stats)+1)
  }

  return(cbind(max_stat, max_stat_index)) #index of last element in first population
}


## Now run the functions over every vector (column) in the matrix ##

# Load the library "snow"" to parallelize the computation using all available cores of the local machine
library(snow)
# Open up sockets to the 4 local processor cores (but only if you have 4 cores...mwahaha)
## Note: only enter "localhost" = to the number of cores on your processor
cl = makeSOCKcluster(c("localhost","localhost","localhost","localhost"))

# Retrieve max test statistics by applying the function to each matrix vector
## Note: expect to wait about 45 mins on a 2.7ghz quad core processor ##
max_stats_anova_mc = parApply(cl, rs, 2, anova_tests)
max_stats_ttest_mc = parApply(cl, rs, 2, t_tests)

```


## Step 3: Create Lambda distribution CDF, PDF and critical values from simulated Lamda Distribution
Now that we have achieved the simulation of the Lamda Distribution we must assemble a CDF and take quantiles of the PDF at different alpha critical value levels.

Our chosen alpha levels:
$$\alpha=0.10, 0.05, 0.025, 0.01, 0.001$$

```{r, tidy = TRUE}
# CDF and plotted PDF using max values Lamda Distribution from ANOVA and t-tests
lambdaCDF_anova = ecdf(max_stats_anova_mc[1,])
plot(density(max_stats_anova_mc[1,]), main = "Lambda PDF", xlab = "Critical Value" , ylab = "Density of Observations")
plot(lambdaCDF_anova, main = "Lambda CDF", xlab = "Critical Value" , ylab = "Accumulated Area")

#t-test is same same
lambdaCDF_ttest = ecdf(max_stats_ttest_mc[1,])
plot(density(max_stats_ttest_mc[1,]), main = "Lambda PDF", xlab = "Critical Value" , ylab = "Density of Observations")
plot(lambdaCDF_ttest, main = "Lambda CDF", xlab = "Critical Value" , ylab = "Accumulated Area")


# Spare P-value function to create P-values from our underlying Lambda Distribution at any specified critical value
## Note: This function is not used in subsequent code, it is just created to provide quick intuition on associated p-values with any critical value if the user wishes to get a better sense of the distribution areas.
lambdaPVal = function (x){
  return(1-lambdaCDF(x))
} 

# Return appropriate critical values of our Lambda Distribution to later test our F-Statistics on live data against 
lambdaQTSd_anova = quantile(max_stats_anova_mc[1,], probs = c(0.90, 0.95, 0.975, 0.99, 0.999), na.rm = FALSE, names = FALSE, type = 1)
lambdaQTSd_ttest = quantile(max_stats_ttest_mc[1,], probs = c(0.90, 0.95, 0.975, 0.99, 0.999), na.rm = FALSE, names = FALSE, type = 1)

```

## Step 4: Perform ANOVA and t-tests on treasury data

```{r, tidy = TRUE}
#find biggest changepoint in data using ANOVA and t-tests
max1 = anova_tests(rates$Value)
max1t = t_tests(rates$Value)

#create two subsets from max1, max1t
pop1a = rates$Value[(1:max1[1,2])]
pop1b = rates$Value[(max1[1,2]+1):(length(rates$Value))]

pop1at = rates$Value[(1:max1t[1,2])]
pop1bt = rates$Value[(max1t[1,2]+1):(length(rates$Value))]

###

#perform tests on each new subset
max2 = anova_tests(pop1a)
max3 = anova_tests(pop1b)

max2t = t_tests(pop1at)
max3t = t_tests(pop1bt)

#create four subsets each from max2, max2t, max3, max3t
pop2a = pop1a[1:max2[1,2]]
pop2b = pop1a[(max2[1,2]+1):length(pop1a)]

pop2at = pop1at[1:max2t[1,2]]
pop2bt = pop1at[(max2t[1,2]+1):length(pop1at)]

pop3a = pop1b[1:max3[1,2]]
pop3b = pop1b[(max3[1,2]+1):length(pop1b)]

pop3at = pop1bt[1:max3t[1,2]]
pop3bt = pop1bt[(max3t[1,2]+1):length(pop1bt)]

###

#perform tests on each new subset
max4 = anova_tests(pop2a)
max5 = anova_tests(pop2b)

max4t = t_tests(pop2at)
max5t = t_tests(pop2bt)

max6 = anova_tests(pop3a)
max7 = anova_tests(pop3bt)

max6t = t_tests(pop3at)
max7t = t_tests(pop3bt)

###

#put maxes in vector
max_stats_anova = c(max1[1], max2[1], max3[1], max4[1], max5[1], max6[1], max7[1])
max_stats_ttest = c(max1t[1], max2t[1], max3t[1], max4t[1], max5t[1], max6t[1], max7t[1])

```

## Step 5: Compare test statistics to lambda critical values

```{r, tidy = TRUE}
#compare each stat to lambda critical values
## function that compares max stat to lambda critical value at alpha=.05
## returns true if s is larger than critical value, false otherwise
compare_max_anova = function(max_stat){
  return(max_stat > lambdaQTSd_anova[4])
}
compare_max_ttest = function(max_stat){
  return(max_stat > lambdaQTSd_ttest[4])
}

compare_anova = unlist(lapply(max_stats_anova, compare_max_anova))
compare_ttest = unlist(lapply(max_stats_ttest, compare_max_ttest))


## function that returns the interest rate for corresponding max stat
find_rate = function(maximum, subpop){
  return (subpop[maximum[1,2]])
}

#apply function to all maxes and subsets
r1 = rates$Value[max1[1,2]]
r2 = find_rate(max2, pop1a)
r3 = find_rate(max3, pop1b)
r4 = find_rate(max4, pop2a)
r5 = find_rate(max5, pop2b)
r6 = find_rate(max6, pop3a)
r7 = find_rate(max7, pop3b)
rlist = c(r1, r2, r3, r4, r5, r6, r7)

##changepoints!!
changepts_anova = cbind(compare_anova, max_stats_anova, rlist)
changepts_ttest = cbind(compare_ttest, max_stats_ttest, rlist)

```

## Step 6: Visulaize the change points

```{r, tidy = TRUE}

#plot rates
plot(rates$Date, rates$Value, type="l")
points(changepts_anova[rlist,], col="red", pch=25)


```

